---
title: "Problem Set 4"
author: "Baskaran Viswanathan"
date: "September 22, 2015"
output: html_document
---


# Q1) Price Histograms with Facet and Color


```{r Price Histograms with Facet and Color }
setwd("~/GitHub/Udacity/EDA_Course_Materials/lesson5")
library(ggplot2)
ggplot(diamonds, aes(x = price, fill = factor(cut))) +
    geom_histogram(binwidth = 200) +
    facet_wrap( ~ color) +
    scale_fill_brewer(type = 'qual')
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#Q2) Findings - price vs. x


looks like there is a linear like relationship between ln(price) and x value. and no X values for some ln(price) valuse between where ln(price)= 7 and 8

There are some outliers such as zero X value and postive price value.


#Q3) Correlations


What is the correlation between price and x?
What is the correlation between price and y?
What is the correlation between price and z?
```{r Correlations }
with(diamonds, cor.test(x=price, y = x, method = 'pearson'))

with(diamonds, cor.test(x=price, y = y, method = 'pearson'))

with(diamonds, cor.test(x=price, y = z, method = 'pearson'))

```


#Q4) price vs. depth


Create a simple scatter plot of price vs depth.

```{r Simple Scatter Plot }
ggplot(data = diamonds, aes(x = depth,y = price)) +
  geom_point(alpha= 0.5)
```


#Q5) Adjustments - price vs. depth


Change the code to make the transparency of the
points to be 1/100 of what they are now and mark
the x-axis every 2 units. See the instructor notes
for two hints.

```{r Adjustments - price vs. depth }
ggplot(data = diamonds, aes(x = depth,y = price)) +
  geom_point(alpha= 0.01) +
  scale_x_continuous(breaks = seq(0,80,2))
  
```


#Q6) Typical Depth Range


Based on the scatterplot of depth vs price, most diamonds are between what values of depth? Answer: 60 - 64


#Q7) Correlation - price and depth


what is the correlation of depth and price?

```{r Correlation - price and depth }

with(diamonds, cor.test(x= depth, y = price, method = "pearson"))

```
We can say there is almost no correlation between depth and price since correlation coefficient is close to zero.


#Q8) price vs. carat


Create a scatterplot of price vs carat
and omit the top 1% of price and carat
values.

```{r price vs. carat }

ggplot(data = diamonds,aes(x = carat, y = price)) + 
  xlim(0,quantile(diamonds$carat,0.99)) +
  ylim(0,quantile(diamonds$price,0.99)) +
  geom_point()

```


#Q9) price vs. volume


Create a scatterplot of price vs. volume (x * y * z).
This is a very rough approximation for a diamond's volume.

Create a new variable for volume in the diamonds data frame.
This will be useful in a later exercise.

```{r price vs. volume }

diamonds$volume <- diamonds$x*diamonds$y*diamonds$z

head(diamonds,3)

ggplot(data = diamonds, aes(x = volume, y = price)) +
  geom_point()

# Some outliers with zero volume

sum(diamonds$volume == 0)

diamonds[diamonds$volume == 0,]

``` 
Looks like there is a correlation, however there are few outlier with high volume with low price makes it hard to see the correlation. Also some (20) diamonds have zero volume



# Q10) Correlations on Subsets


What is the correlation of price and volume? Exclude the diamonds that have a volume of 0 or greater than or equal to 800.

```{r Correlations on Subsets }
with(subset(diamonds, (volume > 0) & (volume <= 800)),cor.test(volume,price))
```


# Q11) Adjustments - price vs. volume


Subset the data to exclude diamonds with a volume
greater than or equal to 800. Also, exclude diamonds
with a volume of 0. Adjust the transparency of the
points and add a linear model to the plot. (See the
Instructor Notes or look up the documentation of
geom_smooth() for more details about smoothers.)

We encourage you to think about this next question and
to post your thoughts in the discussion section.

Do you think this would be a useful model to estimate
the price of diamonds? Why or why not?


```{r Adjustments - price vs. volume }

p1 <- ggplot(data = subset(diamonds, (volume > 0) & (volume <= 800)),
       aes(x = volume, y = price)) +
  geom_point() 

# Default smoother
p2 <- p1 + geom_smooth()

# looking at a linear fit,
p3 <- p1 + stat_smooth(method = "lm", formula = y ~ x, size = 1) + coord_cartesian(ylim = c(0,20000))

# Looking at polynimoal functions of order 2
p4 <- p1 + stat_smooth(method = "lm", formula = y ~ poly(x, 2), size = 1) + coord_cartesian(ylim = c(0,20000))

# Looking at polynimoal functions of order 3
p5 <- p1 + stat_smooth(method = "lm", formula = y ~ poly(x, 3), size = 1) + coord_cartesian(ylim = c(0,20000))

library(gridExtra)
## Loading required package: grid

grid.arrange(p2,p3,p4,p5,ncol =2)

```

# Q12) Mean Price by Clarity


Use the function dplyr package
to create a new data frame containing
info on diamonds by clarity.
Name the data frame diamondsByClarity

The data frame should contain the following
variables in this order.
     (1) mean_price
     (2) median_price
     (3) min_price
     (4) max_price
     (5) n

where n is the number of diamonds in each
level of clarity.

```{r Mean Price by Clarity }
diamondsByClarity <- diamonds %.%
  group_by(clarity) %.%
  summarise(mean_price = mean(price), 
            median_price = median(price), 
            min_price = min(price), 
            max_price = max(price), 
            n= n())

diamondsByClarity

``` 


# Q13) Bar Charts of Mean Price


We've created summary data frames with the mean price
by clarity and color. You can run the code in R to
verify what data is in the variables diamonds_mp_by_clarity
and diamonds_mp_by_color.

Your task is to write additional code to create two bar plots
on one output image using the grid.arrange() function from the package
gridExtra.

Note : BARCHARTS http://docs.ggplot2.org/0.9.3/geom_bar.html 

DIFFERENCE BETWEEN BARCHART AND HISTOGRAM http://flowingdata.com/2014/02/27/how-to-read-histograms-and-use-them-in-r/


```{r Bar Charts of Mean Price }
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1  <- ggplot(diamonds_mp_by_clarity, aes(x = clarity, y = mean_price, fill= clarity)) +
  geom_bar(stat = "identity")

p2 <- ggplot(diamonds_mp_by_color, aes(x = color, y = mean_price, fill= color)) +
  geom_bar(stat = "identity")

grid.arrange(p1,p2, ncol =2)

``` 

s12 has the best mean price whereas VVS1 has the worst mean price. However ther wasn't a very big change in other groups

mean price for color increase from D to J. J has the best mean price and D and E has the worst mean price. This is agains the odds.


``` {r Gapminder Revisited}
diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- summarise(diamonds_by_cut, mean_price = mean(price))

ggplot(diamonds_mp_by_cut, aes(x = cut, y = mean_price, fill= cut)) +
  geom_bar(stat = "identity")
  
```
