---
title: "Problem Set 4"
author: "Baskaran Viswanathan"
date: "September 22, 2015"
output: html_document
---


# Q1) Price Histograms with Facet and Color


```{r Price Histograms with Facet and Color }
setwd("~/GitHub/Udacity/Project3/lesson5")
library(ggplot2)
#ggplot(diamonds, aes(x = price, fill = factor(cut))) +
#    geom_histogram(binwidth = 200) +
#    facet_wrap( ~ color) +
#    scale_fill_brewer(type = 'qual')
#use the fill and scale_fill_brewer to color the histo.
ggplot(aes(x=price, fill = cut), data=diamonds) + 
  geom_histogram() +
  facet_wrap(~color) + 
  scale_fill_brewer(type = 'qual') +
  scale_x_log10()
```



#Q2) Price vs. Table Colored by Cut

Notes: Create a scatterplot of diamond price vs. table and color the points by the cut of the diamond. The plot should look something like this. http://i.imgur.com/rQF9jQr.jpg

Questions: 1. Why does geom_point(aes(color = cut)) work here and not in the previous plot? Or Why does fill not work here on the ggplot layer?

2. Got the color by cut to work but does not look like the sample?


```{r Price vs. Table Colored by Cut}
#ggplot(diamonds, aes(x = table, y = price, color = cut)) +
#    geom_point() +
#    scale_color_brewer(type = 'qual') +
#    scale_x_continuous(breaks = seq(50,80,2), lim = c(50,80))

ggplot(aes(x = table, y = price), data = diamonds) +
  geom_point(aes(color = cut)) +
  scale_fill_brewer(type='qual') +
  coord_cartesian(xlim = c(50,80)) +
  scale_x_discrete(breaks = seq(50,80,2))
```

#Q3) Typical Table Value

What is the typical table range for the majority of diamonds of IDEAL Cut?

Between 55 to 57 

What is the typical table range for the majority of diamonds of PREMIUM Cut?

Between 58 to 62
 

#Q4) Price vs. Volume and Diamond Clarity

Notes: Create a scatterplot of diamond price vs. volume (x * y * z) and color the points by the clarity of diamonds. Use scale on the y-axis to take the log10 of price. You should also omit the top 1% of diamond volumes from the plot.

The plot should look something like this. http://i.imgur.com/excUpea.jpg


```{r Price vs. Volume and Diamond Clarity }

#diamonds$volume = diamonds$x * diamonds$y * diamonds$z
#ggplot(diamonds, aes(x = volume, y = price, color = clarity)) +
#    geom_point() +
#    scale_y_log10() +
#    scale_color_brewer(type = 'div') +
#    xlim(0, quantile(diamonds$volume, 0.99))
#add a volume variable

diamonds$diamond_volume <- diamonds$x * diamonds$y * diamonds$z

#plot scatter of price vs volume colored by clarity

myplot <- ggplot(aes(x = diamond_volume, y = price), data = diamonds) +
  geom_point(aes(color = clarity))

myplot

#change the y axis to log 10 and remove top 1% of diamond volume

myplot <- myplot +
  scale_y_log10() +
  coord_cartesian(xlim=c(0,quantile(diamonds$diamond_volume,0.99)))

myplot

#use color brewer tp adjust color scheme

myplot <- myplot +
  scale_color_brewer(type = 'div')

myplot

```

#Q5) Proportion of Friendships Initiated

Notes: Many interesting variables are derived from two or more others. For example, we might wonder how much of a person's network on a service like Facebook the user actively initiated. Two users with the same degree (or number of friends) might be very different if one initiated most of those connections on the service, while the other initiated very few. So it could be useful to consider this proportion of existing friendships that the user initiated. This might be a good predictor of how active a user is compared with their peers, or other traits, such as personality (i.e., is this person an extrovert?).

Your task is to create a new variable called 'prop_initiated' in the Pseudo-Facebook data set. The variable should contain the proportion of friendships that the user initiated.
```{r Proportion of Friendships Initiated }
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
pf$prop_initiated <- pf$friendships_initiated / pf$friend_count

summary(pf$prop_initiated)

```
#Q6) Prop_initiated vs. tenure


Create a simple scatter plot of price vs depth.

```{r Prop_initiated vs. tenure }
pf$year_joined <- floor(2014 - pf$tenure/365)
pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))

ggplot(data = subset(pf, !is.na(year_joined.bucket)), aes(x = tenure, y = prop_initiated, color = year_joined.bucket)) +
    geom_line(stat = 'summary', fun.y = mean)

```


#Q7) Smoothing prop_initiated vs. tenure


```{r Smoothing prop_initiated vs. tenure }
ggplot(data = subset(pf, !is.na(year_joined.bucket)), aes(x = tenure, y = prop_initiated, color = year_joined.bucket)) +
    geom_line(stat = 'summary', fun.y = mean) +
    geom_smooth()

  
```

#Q8) Greatest prop_initiated Group

On Average which group initiated the greatest proportion of its Facebook friendships

People who joined after 2012 

#Q9) Largest Group Mean prop_initiated

```{r}
subs<-subset(pf, year_joined > 2012)
mean(subs$prop_initiated, na.rm = T)
```



#Q10) Price/Carat Binned, Faceted, & Colored


Image http://i.imgur.com/YzbWkHT.jpg 

```{r Price/Carat Binned, Faceted, & Colored }

ggplot(diamonds, aes(x = cut, y = price/carat)) +
    geom_point(aes(color = diamonds$color), alpha = (1/2), position = position_jitter(width = 0.3)) +
    facet_wrap( ~ clarity) +
    scale_color_brewer(type = 'div')

```



#Q11) Gapminder


Notes: The Gapminder website contains over 500 data sets with information about the world's population. Your task is to continue the investigation you did at the end of Problem Set 4 or you can start fresh and choose a different data set from Gapminder.

If you're feeling adventurous or want to try some data munging see if you can find a data set or scrape one from the web.

In your investigation, examine 3 or more variables and create 2-5 plots that make use of the techniques from Lesson 5.

You can find a link to the Gapminder website in the Instructor Notes.

Once you've completed your investigation, create a post in the discussions that includes: 1. the variable(s) you investigated, your observations, and any summary statistics 2. snippets of code that created the plots 3. links to the images of your plots

Copy and paste all of the code that you used for your investigation, and submit it when you are ready.

# Unemployment in the World
```{r Unemployment in the World }
library(ggplot2)
library(tidyr)
library(dplyr)
```
Assigning to the Variable

```{r Assigning to the Variable}
# read 6 files

# female
f_15_24 <- read.csv("indicator_f 15-24 unemploy - Data.csv", header = TRUE, row.names = 1)
f_25_54 <- read.csv("indicator_f 25-54 unemploy - Data.csv", header = TRUE, row.names = 1)
f_above_55 <- read.csv("indicator_f above 55 unemploy - Data.csv", header = TRUE, row.names = 1)

# male
m_15_24 <- read.csv("indicator_m 15-24 unemploy - Data.csv", header = TRUE, row.names = 1)
m_25_54 <- read.csv("indicator_m 25-54 unemploy - Data.csv", header = TRUE, row.names = 1)
m_above_55 <- read.csv("indicator_m above 55 unemploy - Data.csv", header = TRUE, row.names = 1)

```



```{r add new columns (sex,age,country) to datasets for Females and Males}

#add new columns (sex,age,country) to datasets for Females}

f_15_24['sex'] = 'female'
f_25_54['sex'] = 'female'
f_above_55['sex'] = 'female'

f_15_24['age'] = '15-24'
f_25_54['age'] = '25-54'
f_above_55['age'] = 'above 55'

f_15_24['country'] = row.names(f_15_24)
f_25_54['country'] = row.names(f_25_54)
f_above_55['country'] = row.names(f_above_55)

# add new columns (sex,age,country) to datasets for Females}

m_15_24['sex'] = 'male'
m_25_54['sex'] = 'male'
m_above_55['sex'] = 'male'

m_15_24['age'] = '15-24'
m_25_54['age'] = '25-54'
m_above_55['age'] = 'above 55'

m_15_24['country'] = row.names(m_15_24)
m_25_54['country'] = row.names(m_25_54)
m_above_55['country'] = row.names(m_above_55)

# merge datasets
data <- rbind(f_15_24, f_25_54, f_above_55, m_15_24, m_25_54, m_above_55)

#making tidydata 
t_data <- gather(data, year, total, -sex, -age, -country)

#correct year names
t_data['year'] <- extract_numeric(t_data$year)

```

# First look at tidy_data
```{r First look at tidy_data}

head(t_data)

dim(t_data)

summary(t_data)
```

# Unemployment rate according to gender

Notes: Unemployment quantiles and means for male and female are alomost the same but female boxplot has much higher outliers even over 50% of unemployment.

```{r Unemployment rate according to gender}
ggplot(aes(x = sex, y = total),
       data = subset(t_data, !is.na(total))) + geom_boxplot() +
    stat_summary(fun.y = mean, geom = 'point', shape = 2)

```

# Unemployment mean rate according to gender

Notes: Mean graphs for 1980 - 2005 show that in average female have higher or equal to male unemployment rate.


``` {r Unemployment mean rate according to gender}
ggplot(aes(x = year, y = total),
       data = subset(t_data, !is.na(total))) + 
    geom_line(aes(color = sex), stat = 'summary', fun.y = mean)
```

#Unemployment rate of each country yearly by age 

```{r Unemployment rate of each country yearly by age}
t_data$age <- factor(t_data$age)
t_data$age <- ordered(t_data$age, levels = c('15-24','25-54', 'above 55'))

ggplot(t_data, aes(x = year, y = total)) +
    geom_point(aes(color = t_data$age), alpha = (1/2), position = position_jitter(width = 0.4)) +   
    facet_grid( ~ country) +    
    scale_color_brewer(type = 'qual')
```
# Unemployment analysis by age

Notes: Unemployment is higher in younger age.

```{r Unemployment analysis by age }
ggplot(t_data, aes(x = year, y = total, color = age)) +
    geom_point() +
    scale_color_brewer(type = 'qual')
``` 

